GPU-friendly Stroke Expansion

### 2024-06-28 paper.tex
Rework unrolled recursion section
Explain the unrolled recursion more extensively, and move the discussion into a paragraph of the algorithm section, out of being its own subsection (and not having great flow).

The commit also contains some small fixes found on a closer read.

Correctness is another challenging aspect of stroke rendering. We propose that the correctness of stroke outlines be divided into \emph{weak correctness} and \emph{strong correctness.} We define strong correctness as the computation of the outline of a line swept along the segment, maintaining normal orientation, combined with stroke caps and joins. Weak correctness, by contrast, only requires the parallel curves (also known as ``offset curves'') of path segments, combined with caps and the outer contours of joins. Both \citet{Nehab2020} and \citet{Kilgard2020} provide useful definitions of strongly correct stroke rendering, and Nehab in particular describes how to achieve it in stroke expansion. Briefly, when the path curvature exceeds the reciprocal of the stroke half-width, \emph{evolute} segments and the \emph{inner contours of joins} are required in addition to the parallel curves. The technique described in this paper produces strongly correct outlines according to this definition. Examples of weakly and strongly correct rendering are shown in Figure~\ref{fig:weak_strong}.

@@

The parallel curve of the Euler spiral (also known as ``clothoid'') was characterized by \citet{Wieleitner1907} well over a hundred years ago, and has a straightforward closed-form Cesàro representation, curvature as a function of arc length, of the form $\kappa(s) = c_0{s - s_0}^{-1/2} + c_1$. Euler spirals as an intermediate curve representation for computing parallel curves was proposed in \citet{Levien2021}, which also expands on the Wieleitner reference (including an English translation) and Cesàro equation.

@@

The subdivision density for the parallel curve of an Euler spiral, normalized so that its inflection point is at $-1$ and the cusp of the parallel curve is at 1, is simply $1 - \sqrt{|1-s^2|}$. This function is plotted in Figure~\ref{fig:subdiv_density}.

@@

\subsection{Unrolled recursion}

The entire recursion stack for adaptive subdivision can be represented in two words: $dt$ and $t0\_u$ (the latter being an unsigned integer). The range is $dt \cdot t0\_u .. dt \cdot (t0\_u + 1)$. Initial values of $1$ and $0$ represent the range $(0, 1)$. Pushing the stack, subdividing the range in half, is represented by halving $dt$ and doubling $t0\_u$, which leaves the start of the range invariant but its size halved. After accepting an approximation, the next range is determined by incrementing $t0\_u$. At this point, the decision of whether to pop the stack or continue is dependent on whether $t0\_u$ is even or odd, respectively. If even, popping the stack is accomplished by doubling $dt$ and halving $t0\_u$, and this is repeated until $t0\_u$ becomes odd. The repeated stack popping can be accomplished without iteraction by using the ``count trailing zeros'' intrinsic to determine the number of stack pops $n$, then multiplying $dt$ by $2^n$, and shifting $t0\_u$ right by $n$ bits.

@@

>>>>

Correctness is another challenging aspect of stroke rendering. We propose that the correctness of stroke outlines be divided into \emph{weak correctness} and \emph{strong correctness.} We define strong correctness as the computation of the outline of a line swept along the segment, maintaining normal orientation, combined with stroke caps and joins. Weak correctness, by contrast, only requires the parallel curves (also known as ``offset curves'') of path segments, combined with caps and the outer contours of joins. Both \citet{Nehab2020} and \citet{Kilgard2020} provide useful definitions of strongly correct stroke rendering, and Nehab in particular describes how to achieve it in stroke expansion. Briefly, when the path curvature exceeds the reciprocal of the stroke half-width, \emph{evolute} segments are required in addition to the parallel curves, as well as \emph{inner contours of joins} when segments are short. The technique described in this paper produces strongly correct outlines according to this definition. Examples of weakly and strongly correct rendering are shown in Figure~\ref{fig:weak_strong}.

@@

The parallel curve of the Euler spiral (also known as ``clothoid'') was characterized by \citet{Wieleitner1907} well over a hundred years ago, and has a straightforward closed-form Cesàro representation, curvature as a function of arc length, of the form $\kappa(s) = c_0(s - s_0)^{-1/2} + c_1$. Euler spirals as an intermediate curve representation for computing parallel curves was proposed in \citet{Levien2021}, which also expands on the Wieleitner reference (including an English translation) and Cesàro equation.

@@

The subdivision density for the parallel curve of an Euler spiral, normalized so that its inflection point is at $-1$ and the cusp of the parallel curve is at 1, is simply $\sqrt{|1-s^2|}$. This function is plotted in Figure~\ref{fig:subdiv_density}.

@@

@@

The adaptive subdivision in lowering cubic Béziers to Euler spiral segments is traditionally accomplished through recursion; in the case of subdivision there is a recursive call for each of the two subdivisions of the range. As the execution model of compute shaders does not support recursion, we \emph{unroll} recursion into an iterative loop. The cookbook technique for such unrolling is to store the stack explicitly as an array, which is potentially expensive (it increases pressure on registers). We exploit the specific nature of this recursion, encoding the entire state of the stack into two scalar values: the size of the range $dt$, and the scaled start of the range $t0\_u$, so that the range is $t0\_u \cdot dt .. (t0\_u + 1) \cdot dt$. Initial values are $1$ and $0$ respectively. Pushing the call stack is represented by halving $dt$ and doubling $t0\_u$, which preserves the start of the range but halves the size. Accepting an approximation (a leaf in the call stack) is represented by incrementing $t0\_u$. At this point, the decision of whether to pop the stack or continue is dependent on whether $t0\_u$ is even or odd, respectively. If even, popping the stack is accomplished by doubling $dt$ and halving $t0\_u$, and this is repeated until $t0\_u$ becomes odd. The repeated stack popping can be accomplished without iteraction by using the ``count trailing zeros'' intrinsic to determine the number of stack pops $n$, then multiplying $dt$ by $2^n$, and shifting $t0\_u$ right by $n$ bits.
